### 2020-01-22 README.Rmd
Add some detail to README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Using encounter data to improve capture-recapture abundance estimates"
```

# secmr

## A research compendium for: "Using encounter data to improve capture-recapture abundance estimates"

[![Travis build status](https://travis-ci.org/mbjoseph/secmr.svg?branch=master)](https://travis-ci.org/mbjoseph/secmr)

<!-- This repository contains the data and code for our paper: -->

<!-- > Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx> -->

Our pre-print is online here:

> Joseph MB, Knapp RA, (2020). _`r Title`_. bioRxiv, Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Joseph MB, Knapp RA, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: </archive/master.zip>

Or you can install this compendium as an R package, secmr, from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("mbjoseph/secmr")
```

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.